Gravitational field calculation in spherical coordinates using variable densities in depth

### 2018-09-20 manuscript/manuscript.tex
In the middle of the density based discretization

Variations in the density function of an arbitrary tesseroid introduce
a new type of error into the numerical approximation: only a few point
masses would not reproduce the effect of the density variations.
The modified version of the adaptive discretization algorithm may help
in reducing this kind of error, but as it only takes into account the
dimensions of the tesseroid and the distance to the computation point,
it's not well suited to fully perform this task.

In order to overcome this, we developed a complementary discretization
algorithm that takes into account the variation of the density inside
the tesseroid.
Basically, it subdivides any tesseroid on the radial direction on the
depth at which the \emph{maximum density variation} takes place, only
if a certain discretization inequation holds.

@@

Lets assume an arbitrary \emph{original} tesseroid with boundary radii
$R_1$ and $R_2$ and a density function $\rho(r')$, whose radial
dimension is $L_r = R_2 - R_1$.

First we need to normalise the density function such as its minimum and
maximum will be 0 and 1, respectively.
We call this normalised density $\rho_n(r')$ and we calculate it as
follows:

\begin{equation}
    \rho_n(r') =
    \frac{\rho(r') - \rho_\text{min}}{\rho_\text{max} - \rho_\text{min}},
\end{equation}
\noindent where
\begin{equation}
    \rho_\text{min} = \text{min}\{ \rho(r') \}, \quad
    \rho_\text{max} = \text{max}\{ \rho(r') \} \quad
    \forall \, r' \in [R_1, R_2].
\end{equation}

\noindent Because the algorithm will be iteratively applied for every
subdividing tesseroid, we'd like to note that this normalised density
function will be the same for every subdividing tesseroid.
In case that the density function is constant, both the maximum and
minimum densities are equal and the normalised function could not
be defined, so the density-based discretization algorithm won't be
applied.

Secondly, we must define a linear function $\rho_l(r')$ that assumes
the same values as the normalised density $\rho_n(r')$ on the boundary
radii $R_1$ and $R_2$:

\begin{equation}
    \rho_l(r') =
    \frac{ \rho_n(R_2) - \rho_n(R_1) }{ R_2 - R_1 } (r' - R_1) + \rho_n(R_1),
    \label{eq:density-reference-line}
\end{equation}

\noindent and compute the absolute difference between this linear
function and the normalised density:

Finally, if the following inequation holds, the tesseroid will be divided on
the radius $R_\text{max}$ at which the maximum absolute difference takes
place:

\begin{equation}
    \text{max}\{ \Delta \rho(r') \} \frac{L_r^\text{sm}}{L_r} > \delta,
    \label{eq:delta-density}
\end{equation}

>>>>

The numerical integration of an arbitrary density function introduces
a new type of problem:
the integration error from using only a few nodes to discretize the density function.
The adaptive discretization may help
to reduce this kind of error.
However, it does not take into account the density function
and hence it is not well suited to fully perform this task.

We have developed a complementary discretization
algorithm that takes into account the variations of the density function.
This density-based discretization happens prior to the adaptive discretization described
in the previous section.
In short, the algorithm divides the tesseroid along the radial dimension at the
depths at which the \emph{maximum density variations} take place.

Consider an \emph{original} tesseroid with density given by the function $\rho(r')$.
Before the density-based discretization starts,
we normalise the density function to the range $[0, 1]$ as follows

\begin{equation}
    \rho_n(r') =
    \frac{\rho(r') - \rho_\text{min}}{\rho_\text{max} - \rho_\text{min}},
\end{equation}

\noindent in which $\rho_\text{min}$ and $\rho_\text{max}$ are the minimum and maximum
density values inside the tesseroid boundaries.
We emphasize that this normalised density function will not be modified throughout the
algorithm.
In case the density function is constant, both maximum and minimum densities will be
equal and the density-based discretization algorithm will not be applied.

The algorithm is comprised of the following steps:

\textit{Step 1}:
Define a linear function $\rho_l(r')$ that assumes
the same values as the normalised density $\rho_n(r')$ on the boundaries of the
tesseroid ($r_1$ and $r_2$):

\begin{equation}
    \rho_l(r') =
    \frac{ \rho_n(r_2) - \rho_n(r_1) }{ r_2 - r_1 } (r' - r_1) + \rho_n(r_1),
    \label{eq:density-reference-line}
\end{equation}

\textit{Step 2}:
Evaluate the normalized and linear density functions on a range of $N$ radii between
$r_1$ and $r_2$.
We have opted for $N = 101$ but the specific value of $N$ is not critical to the
algorithm.

\textit{Step 3}:
Compute the absolute difference between the values of the linear and normalised density
functions:

\textit{Step 4}:
If the following inequality holds, the tesseroid will not be divided:

\begin{equation}
    \text{max}\{ \Delta \rho(r') \} \frac{L_r}{L_r^\text{orig}} \le \delta,
    \label{eq:delta-density}
\end{equation}

\noindent
in which $L_r^\text{orig}$ is the radial dimension of the original tesseroid and
$\delta$ is a positive constant henceforth called the \textit{delta ratio}.

on
the radius $R_\text{max}$ at which the maximum absolute difference takes
place:

@@
